Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-07-14 Documents/M1/dgM1Curvilinear.tex
Keep updates.

The two terms of electron-neutrino absorption on nucleons are
\begin{align}
\chi_{nucleon}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} \eta_{np}(g_V^2+3g_A^2)[1-F_e(\varepsilon+Q)](\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonabsorption} ,\\
\eta_{nucleon}(\epsilonNu,\bU) &= \dfrac{G^2}{\pi}\eta_{pn}(g_V^2+3g_A^2)F_e(\varepsilon+Q)(\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonemissivity} ,
\end{align}
and the two terms of electron-neutrino absorption on nuclear are
\begin{align}
\chi_{nuclear}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} n_A e^{(\mu_n -\mu_p -Q')/k_B T} g^2_A \dfrac{2}{7} N_p(Z) N_h(N)[1-F_e(\varepsilon+Q')](\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearabsorption} ,\\
\eta_{nuclear}(\epsilonNu,\bU) &=\dfrac{G^2}{\pi} n_A g^2_A \dfrac{2}{7} N_p(Z) N_h(N)F_e(\varepsilon+Q')(\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearemissivity},
\end{align} 
where $g_V = 1$ and $g_A =1.23$ are form factors resulting from virtual strong-interaction processes, $M_e = 0.511$ MeV is the electron rest energy, $Q = 1.2935$ MeV is the difference between the neutron and proton rest energies, $\mu_n$ and $\mu_p$ are the neutron and proton chemical potential, respectively. And $Q'$ is defined by $Q' = \mu_n -\mu_p + \Delta$, where $\Delta=3$MeV is the energy of the $1f_{5/2}$ level above the ground state.

@@

In Eq.~\eqref{eq:nucleonabsorption} and ~\eqref{eq:nucleonemissivity}, $\eta_{np}$ and $\eta_{pn}$ take into account the nucleon final-state blocking and equal to $n_n$ and $n_p$, the nucleon number density, when the nucleons are nondegenerate, respectively.
\begin{align}
n_p & =  \dfrac{X_p \rho}{m_p} \\
n_n & =  \dfrac{X_n \rho}{m_n},
\end{align}
where $X_p$ and $X_n$ are the proton and neutron fractions, respectively.

@@

The factor $G^2$ in Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} is the Fermi constant [$G=(1.015\pm 0.03)\time 10^5 M_p^{-2}$] with the value
\begin{equation}
G^2 = \left[ \dfrac{1}{(\hbar c)^4} \right] G^2 = 5.18 \times 10^{-44} MeV^{-2}cm^2, 
\end{equation}
giving $\chi$ and $\eta$ the required units of $cm^{-1}$.

@@

For nondegenerate, nonrelativistic nucleons, $\eta_{pp}$ and $\eta_{nn}$ are given by
\begin{align}
&\eta_{pp} = n_p \\
&\eta_{nn} = n_n.
\end{align}

Expanding the isoenergetic scattering kernel in the first two terms of a Legendre series yields
\begin{align}
R^{0}_{IS}(\omega, \omega, \cos\theta) \approx \dfrac{1}{2} \Phi_{0,IS}(\omega) + \dfrac{3}{2}\Phi_{1,IS}(\omega)\cos\theta
\end{align}

@@

Expanding the neutrino-electron scattering kernels in the first two terms of Legendre series yields
\begin{align}
R^{in/out}_{NES}(\omega, \omega^{\prime}, \cos\theta, \bU) & = \dfrac{1}{2} \sum_{l} (2l+1)\Phi^{in/out}_{l,NES}(\omega,\omega^{\prime},\bU)P_l(\cos\theta) \\
& \approx \dfrac{1}{2}\Phi_{0,NES}^{in/out}(\omega,\omega^{\prime}, \bU) + \dfrac{3}{2}\Phi^{in/out}_{1,NES}(\omega,\omega^{\prime}, \bU)\cos\theta
\end{align}

>>>>

The two terms of electron-neutrino absorption on nucleons are
\begin{align}
\chi_{nucleon}(\epsilonNu,\bU) & = \dfrac{G^2_F}{\pi} \eta_{np}(g_V^2+3g_A^2)[1-F_e(\varepsilon+Q)](\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonabsorption} ,\\
\eta_{nucleon}(\epsilonNu,\bU) &= \dfrac{G^2_F}{\pi}\eta_{pn}(g_V^2+3g_A^2)F_e(\varepsilon+Q)(\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonemissivity} ,
\end{align}
and the two terms of electron-neutrino absorption on nuclear are
\begin{align}
\chi_{nuclear}(\epsilonNu,\bU) & = \dfrac{G^2_F}{\pi} n_A e^{(\mu_n -\mu_p -Q')/k_B T} g^2_A \dfrac{2}{7} N_p(Z) N_h(N)[1-F_e(\varepsilon+Q')](\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearabsorption} ,\\
\eta_{nuclear}(\epsilonNu,\bU) &=\dfrac{G^2_F}{\pi} n_A g^2_A \dfrac{2}{7} N_p(Z) N_h(N)F_e(\varepsilon+Q')(\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearemissivity},
\end{align} 
where $g_V = 1$ and $g_A =1.23$ are form factors resulting from virtual strong-interaction processes, $M_e = 0.511$ MeV is the electron rest energy, $Q = 1.2935$ MeV is the difference between the neutron and proton rest energies, $\mu_n$ and $\mu_p$ are the neutron and proton chemical potential, respectively. And $Q'$ is defined by $Q' = \mu_n -\mu_p + \Delta$, where $\Delta=3$MeV is the energy of the $1f_{5/2}$ level above the ground state.

The factor $G^2_F$ in Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} is the Fermi constant [$G=(1.015\pm 0.03)\time 10^5 M_p^{-2}$] with the value
\begin{equation}
G^2_F = \left[ \dfrac{1}{(\hbar c)^4} \right] G^2 = 5.18 \times 10^{-44} MeV^{-2}cm^2, 
\end{equation}
giving $\chi$ and $\eta$ the required units of $cm^{-1}$.

@@

In Eq.~\eqref{eq:nucleonabsorption} and ~\eqref{eq:nucleonemissivity}, $\eta_{np}$ and $\eta_{pn}$ take into account the nucleon final-state blocking. 
\begin{align}
\eta_{pn} = \dfrac{2}{h^3 c^3} \int d^3 p F_p (E)[1-F_n(E)] = (n_n - n_p) \dfrac{1}{e^{(\mu_n^0 - \mu_p^0)/k_bT}-1},\\
\eta_{np} = \dfrac{2}{h^3 c^3} \int d^3 p F_n (E)[1-F_p(E)] = (n_p - n_n) \dfrac{1}{e^{(\mu_p^0 - \mu_n^0)/k_bT}-1}.
\end{align}

The factors $\eta_{np}$ and $\eta_{pn}$ equal to $n_n$ and $n_p$, the nucleon number density, when the nucleons are non-degenerate, respectively.
\begin{align}
n_p & =  \dfrac{X_p \rho}{m_p} \\
n_n & =  \dfrac{X_n \rho}{m_n},
\end{align}
where $X_p$ and $X_n$ are the proton and neutron fractions, respectively.

@@

@@

The factors $\eta_{pp}$ and $\eta_{nn}$ account for the nucleon final-state blocking,
\begin{align}
\eta_{pp} = T \dfrac{\partial n_p}{\partial \mu_p} = n_p \dfrac{3T/2\epsilon^p_F}{[1+(3T/2\epsilon^p_F)^2]^{1/2}}, \\
\eta_{nn} = T \dfrac{\partial n_n}{\partial \mu_n}=n_n \dfrac{3T/2\epsilon^n_F}{[1+(3T/2\epsilon^n_F)^2]^{1/2}},
\end{align}
where $\epsilon^p_F$ and $\epsilon^n_F$ are the proton and neutron Fermi energies, respectively,
\begin{align}
\epsilon^p_F = \dfrac{h^2}{8 \pi^2 m_B}(3\pi^2n_p)^{2/3},\\
\epsilon^n_F = \dfrac{h^2}{8\pi^2m_B}(3\pi^2n_n)^{2/3}.
\end{align}
For nondegenerate, nonrelativistic nucleons, $\eta_{pp}$ and $\eta_{nn}$ are given by
\begin{align}
&\eta_{pp} = n_p \\
&\eta_{nn} = n_n.
\end{align}

Expanding the isoenergetic scattering kernel in the first two terms of a Legendre series yields
\begin{align}
R_{IS}(\omega, \omega, \cos\theta) \approx \dfrac{1}{2} \Phi_{0,IS}(\omega) + \dfrac{3}{2}\Phi_{1,IS}(\omega)\cos\theta
\end{align}

@@

Expanding the neutrino-electron scattering kernels in the first two terms of Legendre series yields
\begin{align}
R^{in/out}_{NES}(\omega, \omega^{\prime}, \cos\theta, \bU) & = \sum_{l=0}^{1} \dfrac{ 2l+1}{2}\Phi^{in/out}_{l,NES}(\omega,\omega^{\prime},\bU)P_l(\cos\theta) \\
& \approx \dfrac{1}{2}\Phi_{0,NES}^{in/out}(\omega,\omega^{\prime}, \bU) + \dfrac{3}{2}\Phi^{in/out}_{1,NES}(\omega,\omega^{\prime}, \bU)\cos\theta
\end{align}
